---
title: "Honors Thesis"
format: pdf
editor: visual
---

## Libraries

```{r}
library(tidyverse)
library(readxl)
library(lubridate)
```

## Clean QWI Distress Data

```{r}
raw_qwi <- read_csv("~/Downloads/qwi_f5513339c9334a45b599500816c6a08e.csv")

clean_qwi <- raw_qwi %>%
  filter(firmsize %in% c(1, 2, 3, 4, 5)) %>%
  
  mutate(
    Size_Category = case_when(
      firmsize == 1 ~ "Small (0-19 emps)",
      firmsize == 2 ~ "Small-Med (20-49 emps)",
      firmsize == 3 ~ "Medium (50-249 emps)",
      firmsize == 4 ~ "Med-Large (250-499 emps)",
      firmsize == 5 ~ "Large (500+ emps)"
    ),
    
    Year_Quarter = paste0(year, "-Q", quarter),
    
    # THE INDUSTRY FIX: Tell R that "31-33" means "33" so it matches the crosswalk
    NAICS_Code = ifelse(industry == "31-33", "33", as.character(industry))
  ) %>%
  
  # THE METRIC FIX: Use 'FrmJbLs' (Firm Job Loss) instead of 'Sep'
  dplyr::select(Year_Quarter, year, quarter, NAICS_Code, Size_Category, Emp, FrmJbLs)
```

## Clean Concordance Code Data

```{r}
raw_crosswalk <- read_excel("~/Downloads/impconcord25.xlsx", col_types = "text")

clean_crosswalk <- raw_crosswalk %>%
  mutate(
    # Make code the international 6 digit standard
    HTS_6 = substr(commodity, 1, 6),
    
    # Match the 6 digit to the 2 digit QWI sectors 
    NAICS_Code = substr(naics, 1, 2)) %>%
  distinct(HTS_6, NAICS_Code)
```

## Combine and Clean Tariff Shock Data

```{r}
library(tidyverse)

tariff_2018 <- read_excel("~/Downloads/tariff_data_2018/tariff_database_2018.xlsx", col_types = "text") %>% 
  mutate(year = 2018, mfn_ad_val_rate = as.numeric(mfn_ad_val_rate))

tariff_2019 <- read_excel("~/Downloads/tariff_data_2019/2019_Tariff_Database_v11.xlsx", col_types = "text") %>% 
  mutate(year = 2019, mfn_ad_val_rate = as.numeric(mfn_ad_val_rate))

tariff_2025 <- read_excel("~/Downloads/tariff_data_2025/tariff_database_2025.xlsx", col_types = "text") %>% 
  mutate(year = 2025, mfn_ad_val_rate = as.numeric(mfn_ad_val_rate))

tariff_2026 <- read_excel("~/Downloads/tariff_data_2026/tariff_database_2026.xlsx", col_types = "text") %>% 
  mutate(year = 2026, mfn_ad_val_rate = as.numeric(mfn_ad_val_rate))

# Make them one master dataframe
raw_tariffs <- bind_rows(tariff_2018, tariff_2019, tariff_2025, tariff_2026)

# Clean combined tariff data 
clean_tariffs <- raw_tariffs %>%
  mutate(
    HTS_6 = substr(hts8, 1, 6),
    Statutory_Tariff_Rate = mfn_ad_val_rate * 100
  ) %>%
  filter(!is.na(Statutory_Tariff_Rate)) %>%
  
  # Average the 8 digit rates into a single 6 digit rate per year
  group_by(year, HTS_6) %>%
  summarize(
    Statutory_Tariff_Rate = mean(Statutory_Tariff_Rate, na.rm = TRUE),
    .groups = 'drop'
  )

```

## Merge all

```{r}
# Attach Industry codes to the Tariff Data
tariffs_with_industry <- clean_tariffs %>%
  left_join(clean_crosswalk, by = "HTS_6", relationship = "many-to-many") %>%
  filter(!is.na(NAICS_Code))

# Aggregate Tariffs to the Industry-Year level
industry_tariffs <- tariffs_with_industry %>%
  group_by(year, NAICS_Code) %>%
  summarize(
    Avg_Tariff_Rate = mean(Statutory_Tariff_Rate, na.rm = TRUE),
    .groups = 'drop'
  )

# Final Join
final_did_panel <- clean_qwi %>%
  left_join(industry_tariffs, by = c("year", "NAICS_Code")) %>%
  mutate(
    Avg_Tariff_Rate = replace_na(Avg_Tariff_Rate, 0),
    Is_Treated = ifelse(Avg_Tariff_Rate > 0, 1, 0),
    
    # Calculate the Distress Metric (Jobs Destroyed / Total Employment)
    Job_Loss_Rate = FrmJbLs / Emp
  )
```

## Bring is BTOS index data for 2025-2026

```{r}

raw_btos <- read_excel("~/Downloads/Sector by Employment Size Class.xlsx", sheet = "Index Estimates", col_types = "text")

clean_btos <- raw_btos %>%
  # Filter to specific size brackets 
  filter(!is.na(Sector), nchar(Sector) <= 2, Empsize != "XX") %>%
  
  # Rename 'Sector' to match the tariff dataset
  rename(NAICS_Code = Sector) %>%
  
  # Take the date columns and melt them down into two new columns
  pivot_longer(
    cols = starts_with("20"), 
    names_to = "Cycle_Date", 
    values_to = "Distress_Index"
  ) %>%
  
  # Clean up the dates and sizes so we can join to the tariffs
  mutate(
    year = as.numeric(substr(Cycle_Date, 1, 4)),
    
    # Translate the BTOS size codes to readable labels
    Size_Category = case_when(
      Empsize == "A" ~ "Small (1-4 emps)",
      Empsize == "B" ~ "Small (5-9 emps)",
      Empsize %in% c("C", "D") ~ "Medium (10-49 emps)",
      Empsize %in% c("E", "F") ~ "Med-Large (50-249 emps)",
      Empsize == "G" ~ "Large (250+ emps)"
    ),
    Distress_Index = ifelse(Distress_Index == ".", NA, Distress_Index),
    Distress_Index = as.numeric(Distress_Index)
  ) %>%
  
  # Filter out missing data
  filter(!is.na(Distress_Index)) %>%
  
  # Keep only what we need
  dplyr::select(year, Cycle_Date, NAICS_Code, Size_Category, `Option Text`, Distress_Index)

final_btos <- clean_btos %>%
  left_join(industry_tariffs, by = c("year", "NAICS_Code")) %>%
  mutate(
    # If an industry had no tariffs, set it to a 0% rate
    Avg_Tariff_Rate = replace_na(Avg_Tariff_Rate, 0),
    
    # Create the Treatment Dummy for the regression
    Is_Treated = ifelse(Avg_Tariff_Rate > 0, 1, 0)
  ) %>%
  # Focus specifically on the "Future demand"
  filter(`Option Text` == "Future demand")
```

## Linear Model for 2018-2019

```{r}
model_1 <- lm(Job_Loss_Rate ~ Avg_Tariff_Rate + Size_Category + (Avg_Tariff_Rate * Size_Category), data = final_did_panel)
summary(model_1)
```

In 2018-2019, small-medium and medium businesses were hit the hardest compared to large businesses and show high statistical significance with p-values close to 0. Surprisingly, the most small size of businesses did not show statistical significance in getting hit by the tariffs compared to large businesses, but sometimes these businesses are so niche that it isolates them from tariff influenced industries.

## Linear Model for 2025-2026

```{r}
model_2 <- lm(Distress_Index ~ Avg_Tariff_Rate + Size_Category + (Avg_Tariff_Rate * Size_Category), data = final_btos)
summary(model_2)
```

It seems for 2025-206, businesses are being hurt from tariffs across the board. It is not specifically hitting small businesses worse according to business survey responses. Medium is the closest to seeing statistical significance though, which is similar to 2018-2019 and could be a sign that trends will be the same. It is also important to note that even if small businesses have the same impact as larger businesses, smaller businesses are less equipped to deal with the negative impact.

The 2018 data does also show a lagged effect, so we may see this impact on the 2025-2026 more in the future.

```{r}
library(lmtest)
library(sandwich)
library(car)

par(mfrow = c(2, 2)) 
plot(model_1) 
par(mfrow = c(1, 1)) 


robust_model_1 <- coeftest(model_1, vcov = vcovHC(model_1, type = "HC1"))
print(robust_model_1)


vif(model_1)
```

```{r}

model_glm <- glm(Job_Loss_Rate ~ Avg_Tariff_Rate + Size_Category + (Avg_Tariff_Rate * Size_Category), 
                 family = quasibinomial(link = "logit"), 
                 data = final_did_panel)

summary(model_glm)

par(mfrow = c(2, 2)) 
plot(model_glm) 
par(mfrow = c(1, 1)) 

robust_model_glm <- coeftest(model_glm, vcov = vcovHC(model_glm, type = "HC1"))
print(robust_model_glm)


vif(model_glm)
```

The initial OLS linear model showed that tariffs increased job destruction more in small-medium businesses than the smallest one. However, diagnostics showed the though there was not a multicollinearity problem, there was heteroskedasticity. To correct this, I tried a GLM model, and micro small companies showed the biggest hit with tariffs followed by small-medium ones.
